A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)

### 2024-12-15 manuscript/manuscript.qmd
minor formatting

The focus of ENCORE is to maximize potential learnings on when RWE studies can or cannot yield similar results compared to RCTs. To that end, the emphasis of the project is on trials of therapies for the most common cancers and/or cancers for which there has been substantial therapeutic development in recent years. After careful review and exchange with clinical and regulatory experts, four cancer indications were identified including lung cancer, breast cancer, colorectal cancer and multiple myeloma. For each cancer we aim to conduct three trial emulations which will be implemented using multiple databases accessible for the scope of this project (i.e. the total number of emulations will equal 12 trials x *n* databases which are found fit-for-purpose for each trial).

@@

**Statistical power**. Causal analyses of observational data may not have the same pre-requisites in terms of formal hypothesis testing and statistical power than RCTs since the number of 'recruited' patients is given and cannot be influenced.[@hernan2022causal] For this project, however, statistical power is a critical to assess the feasibility that a comparison of agreement metrics of estimated parameters between RCT and RWD with sufficient precision can be made. Since the main outcome of interest is defined as time to all-cause mortality (OS), the estimation of the statistical power is driven by the number of events rather than the number of patients. To assess if the unstratified number of events is sufficient such that a significant difference can be detected based on the original RCT-reported hazard ratio (HR), the statistical power $\beta$ will be estimated using Schoenfeld's sample-size formula for the proportional-hazards regression model.[@schoenfeld1981asymptotic; @schoenfeld1983sample; @gsDesign]

@@

This issue may be even augmented with more complex treatment protocols like combination regimens (e.g., CheckMate9LA) as compared to monotherapies (CheckMate057) since the ascertainment of the exposure needs to happen over a pre-defined time window which is often difficult to calibrate in RWD. As a result, exposure misclassification (due to too narrow ascertainment windows) and selection bias (due to too long ascertainment windows) are common trade-offs in such scenarios. Alternative analytic approaches which target a per-protocol estimand, such as the clone-censor-weight design[@gaber2024mystifying], may be viable options if the ITT estimand cannot be estimated due to the aforementioned parameters and provided that necessary covariate measurements are available to account for the artificial censoring introduced with these methods.

>>>>

The focus of ENCORE is to maximize potential learnings on when RWE studies can or cannot yield similar results compared to RCTs. To that end, the emphasis of the project is on trials of therapies for the most common cancers and/or cancers for which there has been substantial therapeutic development in recent years. After careful review and exchange with clinical and regulatory experts, four cancer indications were identified including lung cancer, breast cancer, colorectal cancer and multiple myeloma. For each cancer we aim to conduct three trial emulations which will be implemented using multiple databases accessible for the scope of this project (i.e., the total number of emulations will equal 12 trials x *n* databases which are found fit-for-purpose for each trial).

@@

#### **Statistical power**

Causal analyses of observational data may not have the same pre-requisites in terms of formal hypothesis testing and statistical power than RCTs since the number of 'recruited' patients is given and cannot be influenced.[@hernan2022causal] For this project, however, statistical power is a critical to assess the feasibility that a comparison of agreement metrics of estimated parameters between RCT and RWD with sufficient precision can be made. Since the main outcome of interest is defined as time to all-cause mortality (OS), the estimation of the statistical power is driven by the number of events rather than the number of patients. To assess if the unstratified number of events is sufficient such that a significant difference can be detected based on the original RCT-reported hazard ratio (HR), the statistical power $\beta$ will be estimated using Schoenfeld's sample-size formula for the proportional-hazards regression model.[@schoenfeld1981asymptotic; @schoenfeld1983sample; @gsDesign]

@@

This issue may be even augmented with more complex treatment protocols like combination regimens (e.g., CheckMate9LA) as compared to monotherapies (e.g., CheckMate057) since the ascertainment of the exposure needs to happen over a pre-defined time window which is often difficult to calibrate in RWD. As a result, exposure misclassification (due to too narrow ascertainment windows) and selection bias (due to too long ascertainment windows) are common trade-offs in such scenarios. Alternative analytic approaches which target a per-protocol estimand, such as the clone-censor-weight design[@gaber2024mystifying], may be viable options if the ITT estimand cannot be estimated due to the aforementioned parameters and provided that necessary covariate measurements are available to account for the artificial censoring introduced with these methods.
